---
title: "Beyoncé and Taylor Swift Lyrics"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(skimr)
library(tidytext)  # for working with text data
library(lubridate) # for working with dates
library(scales)    # for scales on plots 
library(patchwork) # for arranging plots
```

The data come from [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-29/readme.md). TidyTuesday is a weekly social data project for the R community. Read more about TidyTuesday [here](https://github.com/rfordatascience/tidytuesday) and see people's contributions on Twitter under the [#tidytuesday hashtag](https://twitter.com/search?q=tidytuesday&src=typed_query). 

```{r load-data, message = FALSE}
beyonce_lyrics <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv')
tswift_lyrics  <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv')
sales          <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv')
charts         <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/charts.csv')
```

## Data overview

```{r skim-sales}
skim(sales)
```

```{r skim-charts}
skim(charts)
```

## Sales

#### Sales by artist and country over time

```{r}
sales %>%
  filter(country == "UK" | country == "US") %>%
  ggplot(aes(y = sales, x = year(mdy(released)), color = artist)) +
  geom_point() +
  facet_wrap(~country, nrow = 2) +
  theme_minimal() +
  scale_y_continuous(labels = label_dollar()) +
  labs(
    x = "Year released",
    y = "Sales (in USD)",
    color = "Artist"
  )
```



## Charts

### Average chart position by artist and country

```{r warning = FALSE, fig.asp = 0.5, out.width = "90%"}
charts %>%
  group_by(artist, chart) %>%
  summarise(avg_chart_pos = mean(as.numeric(chart_position), na.rm = TRUE)) %>%
  ggplot(aes(x = chart, y = avg_chart_pos)) +
  geom_col() +
  facet_wrap(~artist)
```

### Chart position by artist and country

```{r warning = FALSE, fig.asp = 0.5, out.width = "90%"}
ggplot(charts, aes(y = chart, x = as.numeric(chart_position), color = artist)) +
  geom_boxplot()
```

What's that one Taylor Swift album that did so badly in the UK?

```{r warning = FALSE}
charts %>%
  filter(as.numeric(chart_position) > 80) %>%
  select(artist, title)
```


## Text analysis

### Common words in Beyonce and Taylor Swift lyrics

```{r warning = FALSE, fig.asp = 0.5, out.width = "90%"}
p1 <- beyonce_lyrics %>%
  unnest_tokens(word, line) %>%
  anti_join(get_stopwords()) %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(y = fct_reorder(word, n), x = n)) +
  geom_col() +
  labs(title = "Beyonce")

p2 <- tswift_lyrics %>%
  unnest_tokens(word, Lyrics) %>%
  anti_join(get_stopwords()) %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(y = fct_reorder(word, n), x = n)) +
  geom_col() +
  labs(title = "Taylor Swift")

p1 + p2
```



