---
title: "Himalayan climbing expeditions"
author: "Mine Çetinkaya-Rundel"
date: "24 September 2020"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(skimr)
```

The data come from [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-22/readme.md). TidyTuesday is a weekly social data project for the R community. Read more about TidyTuesday [here](https://github.com/rfordatascience/tidytuesday) and see people's contributions on Twitter under the [#tidytuesday hashtag](https://twitter.com/search?q=tidytuesday&src=typed_query). 

```{r load-data, message = FALSE}
members     <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-22/members.csv")
expeditions <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-22/expeditions.csv")
peaks       <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-22/peaks.csv")
```

## Data overview

### Members

```{r}
skim(members)
```

### Peaks

```{r}
skim(peaks)
```


## Popular peaks

```{r}
members %>%
  count(peak_name) %>%
  arrange(desc(n)) %>%
  mutate(proportion = n / sum(n))
```


## Hired and died

```{r}
members %>%
  count(hired, died) %>%
  group_by(hired) %>%
  mutate(proportion = n / sum(n))
```

```{r}
ggplot(members, aes(x = hired, fill = died)) +
  geom_bar(position = "fill")
```

Go to RStudio Cloud and try it yourself in the project titled `Code Along 01 - Himalayan Climbing Expeditions`.