---
title: "UK Baby Names"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r chunk-options, include = FALSE}
# what is this?
knitr::opts_chunk$set(fig.width = 6, fig.asp = 0.618, dpi = 300, out.width = "80%")
```

```{r load-packages, message = FALSE}
library(tidyverse)
library(ukbabynames)
```

## Data

We'll work with the [**ukbabynames**](https://github.com/mine-cetinkaya-rundel/ukbabynames) package, which has two data sets:

- `ukbabynames`: Full baby name 1996-2017 for the United Kingdom from the Office of National Statistics
- `rankings`: A longitudinal dataset containing the top-100 baby girl and top-100 baby boy names over the period 1904 to 1994.

Let's save these data as tibble to make them easier to work with.

```{r make-tibbles}
ukbabynames <- as_tibble(ukbabynames)
rankings <- as_tibble(rankings)
```

Using this dataset, we can, for example, take a look at how popular a particular name has been over the years.

## Most popular first names in IDS

The most popular first names in IDS are as follows:

```
first_name     n
Alex           4
James          3
Sophie         3
Tom            3
```

Let's see how these names fared over time in the UK.

```{r ids-names-rank}
popular_IDS_names <- c("Alex", "James", "Sophie", "Tom")

ukbabynames %>%
  filter(name %in% popular_IDS_names) %>%
  ggplot(aes(x = year, y = rank, color = name)) +
  geom_line() +
  facet_wrap(~sex)
```

What does a higher rank mean? More or less popular? How can we represent this better on the plot?

```{r rank-rep}
ukbabynames %>%
  filter(name %in% popular_IDS_names) %>%
  ggplot(aes(x = year, y = rank, color = name)) +
  geom_line() +
  facet_wrap(~sex) +
  scale_y_reverse()
```

Why does the y-axis go up until 6000? How can we fix this?

```{r fix-y-axis}
ukbabynames %>%
  filter(name %in% popular_IDS_names) %>%
  arrange(desc(rank))

ukbabynames %>%
  filter(name %in% popular_IDS_names) %>%
  filter(!(name == "James" & sex == "F")) %>% # take out single James in females
  ggplot(aes(x = year, y = rank, color = name)) +
  geom_line() +
  facet_wrap(~sex) +
  scale_y_reverse()
```

## Coding style

As we analyse the data further, we'll give pointers on coding style. 

So first, a mini-lesson on coding style!

>"Good coding style is like correct punctuation: you can manage without it, butitsuremakesthingseasiertoread."
>
>Hadley Wickham

We'll use the [Tidyverse style guide](http://style.tidyverse.org/). There's more to it than what we'll cover today, but we'll mention more as we introduce more functionality, and do a recap later in the semester.

### File names and code chunk labels

- Do not use spaces in file names, use `-` or `_` to separate words
- Use all lowercase letters

```{r file-names, eval = FALSE}
# Good
uk-baby-names.csv

# Bad
UK Baby Names.csv
```

## Object names

- Use `_` to separate words in object names
- Use informative but short object names
- Do not reuse object names within an analysis

```{r ukbabynames-2017}
# Good
ukbabynames_2017 <- ukbabynames %>% 
  filter(year == 2017)
```

```{r ukbabynames-2017-bad-names, eval = FALSE}
# Bad
ukbabynames.2017
ukbabynames2
ukbabynames_subset
ukbabynames_subsetted_for_2017
```


## Spacing

- Put a space before and after all infix operators (`=`, `+`, `-`, `<-`, etc.), and when naming arguments in function calls. 
- Always put a space after a comma, and never before (just like in regular English).

```{r alex}
# Good
alex <- ukbabynames %>% 
  group_by(year, sex) %>%
  mutate(prop = n / sum(n)) %>%
  filter(name == "Alex")
```

```{r alex-bad-specing, eval = FALSE}
# Bad
alex<-ukbabynames%>% 
  group_by(year,sex)%>%
  mutate(prop=n/sum( n ))%>%
  filter(name=="Alex")
```

## ggplot

- Always end a line with `+`
- Always indent the next line

```{r alex-plot}
# Good
ggplot(alex, aes(x = year, y = prop, color = sex, group = sex)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Proportion of people named Alex",
    color = "Sex",
    title = "Alex over the years in the UK"
  )
```

```{alex-plot-bad-formatting, eval = FALSE}
# Bad
ggplot(alex,aes(x=year,y = prop,color=sex,group=sex))+geom_line()+
  labs(x = "Year",y = "Proportion of people named Alex",color = "Sex",title = "Alex over the years in the UK")
```

## Long lines

- Limit your code to 80 characters per line. This fits comfortably on a printed page with a reasonably sized font.

- Take advantage of RStudio editor's auto formatting for indentation at line breaks.

```{alex-plot-bad-formatting-long-lines, eval = FALSE}
# Bad
ggplot(alex,aes(x=year,y = prop,color=sex,group=sex))+geom_line()+labs(x = "Year",y = "Proportion of people named Alex",color = "Sex",title = "Alex over the years in the UK")
```
  
## Assignment
  
Use `<-` not `=`

```{r jude}
# Good
jude <- ukbabynames %>% 
  group_by(year, sex) %>%
  mutate(prop = n / sum(n)) %>%
  filter(name == "Jude")
```

```{r jude-bad-assignment, eval = FALSE}
# Bad
taylor = ukbabynames %>% 
  group_by(year, sex) %>%
  mutate(prop = n / sum(n)) %>%
  filter(name == "Taylor")
```
  
## Quotes
  
Use `"`, not `'`, for quoting text. The only exception is when the text already contains double quotes and no single quotes.

```{r jude-plot}
ggplot(jude, aes(x = year, y = prop, color = sex, group = sex)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Proportion of people named Taylor",
    color = "Sex",
    title = 'Hey Jude"'
  )
```

## When was the last person called Gary born

In this dataset, what is the birth year of the youngest person named Gary?

```{r young-gary}
ukbabynames %>%
  filter(name == "Gary") %>%
  arrange(desc(year))
```

## Few more names


```{r}
ukbabynames %>% 
  filter(name == "Gleb") %>%
  ggplot(aes(x = year, y = rank, color = sex, group = sex)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Rank of Gleb",
    color = "Sex",
    title = "For Gleb"
  ) +
  scale_y_reverse()
```
